Geometric variational problems

### 2020-11-07 NSF2020/Description/specificaim2_ver2.tex
SA2 close to done

%\subsection{Specific Aim 2: Functional Materials}
\subsection{Specific Aim 2: Hydrodynamics}
\label{subsec:specific_aim_2}
Specific Aim 1 dealt with the collective material property of amphiphiles which does not involve the movement  of the solvent.
In Specific Aim 2, we include hydrodynamics effects. Hydrodynamic effects are relevant
because the rates of biological functions like fusion, fission, and pore dynamics rely on viscous dissipation \cite{RYHAM20112929}. 
Additionally, coarse-grained and molecular dynamics theories include water either explicitly or implicitly in their models.

@@

The particle velocity and rotation $({\bf v}_i, {\bf w}_i)$ allow us to update the particle position:
\begin{alignat}{3}
\label{Hupdate1}   \frac{\dif {\bf a}_i}{\dif t} = {\bf v}_i,      \quad  \frac{\dif {\bf R}_i}{\dif t}  = A({\bf w}_i).
\end{alignat}
Recall that $A({\bf w}_i)$ is the skew-symmetric matrix with axial vector ${\bf w}_i$.
Also, ${\bf a}_i$ is the particle center the rotation matrix ${\bf R}_i$ defines the orientation.
That way, the particle configurations come from the flow maps $F_i({\bf x},t) = {\bf a}_i(t) + {\bf R_i}(t){\bf x}$ with
$P_i(t) = F_i(P_{0i},t)$ and $f_i(F_i({\bf x},t)) = f_{i0}({\bf x})$ where $P_{i0}$ and $f_{i0}$ are the $i$th reference particle
and material label respectively.

>>>>

%\subsection{Specific Aim 2: Functional Materials}
\subsection{Specific Aim 2: Hydrodynamics}
\label{subsec:specific_aim_2}
Specific Aim 1 dealt with the collective material property of amphiphiles which does not involve the movement  of the solvent.
In Specific Aim 2, we include hydrodynamics effects. Hydrodynamic effects are relevant
because the rates of biological functions like fusion, fission, and pore dynamics rely on viscous dissipation \cite{RYHAM20112929},
and can determine which local equilibria the system tends towards. 
Additionally, coarse-grained and molecular dynamics theories include water either explicitly or implicitly in their models.

@@

The particle velocity and rotation $({\bf v}_i, {\bf w}_i)$ update the particle configurations:
\begin{alignat}{3}
\label{Hupdate1}   \frac{\dif {\bf a}_i}{\dif t} = {\bf v}_i,      \quad  \frac{\dif {\bf R}_i}{\dif t}  = A({\bf w}_i),\quad i = 1, \dots, N.
\end{alignat}
Recall that $A({\bf w}_i)$ is the skew-symmetric matrix with axial vector ${\bf w}_i$.
Also, ${\bf a}_i$ is the particle center the rotation matrix ${\bf R}_i$ defines the orientation.
That way, the particle configurations come from the flow maps $F_i({\bf x},t) = {\bf a}_i(t) + {\bf R_i}(t){\bf x}$ with
$P_i(t) = F_i(P_{0i},t)$ and $f_i(F_i({\bf x},t)) = f_{i0}({\bf x})$ where $P_{i0}$ and $f_{i0}$ are the $i$th reference particle
and material label respectively.

Numerically solving \eqref{St1}--\eqref{HSB2} is a nontrivial matter.
First, we have to carefully map the particle configurations because the domain $\Omega(t)$ and boundary conditions $f_i$ are constantly changing.
The particle dynamics in \eqref{Hupdate1} and in \eqref{SGD} both involve solving
screened Laplace boundary value problems on evolving domains, but differ in that \eqref{Hupdate1} has an additional Stokes flow
generating the dissipation. When there is no background flow ${\bf u}_{\infty} = {\bf 0}$, the system obeys the energy law
\begin{equation}
\frac{\dif}{\dif t} \Phi + \int_{\Omega(t)} \mu |\nabla {\bf u} + \nabla {\bf u}^T|^2 \,dx = 0
\end{equation}
and the particle configurations tend toward a local equilibrium.

A nonzero background flow continually injects energy into the system. We deal with the background flow using
the Power and Miranda representation 
\begin{equation}
\label{PowerMiranda}
{\bf u} = {\bf u}_{\infty} + D\eta + \sum_{i=1}^N S(\cdot, {\bf a}_i) {\bf F}_i + R(\cdot, {\bf a}_i) {\bf G}_i.
\end{equation}
where $S(\cdot, {\bf a}_i)$ and $ R(\cdot, {\bf a}_i)$ are stokeslets and rotlets supported at the particle centers ${\bf a}_i.$
The symbol $D$ denotes the Stokes double layer integral operator (see Specific Aim 3).   
The representation \eqref{PowerMiranda} automatically satisfies \eqref{St1}, \eqref{St2}, \eqref{St4} and \eqref{HSB1}.
We solve for the unknown density $\eta : \partial \Omega \to \mathbb{R}$ by requiring that jump in viscous stress
vanished across $\partial P_i$. This jump condition guarantees that ${\bf u}$ is a rigid motion within each particle,
and by continuity, equals a rigid motion on the particle boundary as required by \eqref{St3}.

We prove a, what to our knowledge is new, result concerning the evaluation of Maxwell-type stresses. 
\begin{proposition}
  \label{prop:recip}
  Suppose that $u = \sum_{i=1}^N u_i$ where $u_i$ is a solution of the screened Laplace equation
  in $\mathbb{R}^3 \setminus P_i$ for $i=1,\dots, N.$ Then 
\begin{equation}
{\bf F}_i = \sum_{j \neq i} \int_{\partial P_i}[\boldsymbol{\sigma}_{ij} + \boldsymbol{\sigma}_{ji}]\boldsymbol{\nu}\,\dif S,\quad
{\bf G}_i = \sum_{j \neq i} \int_{\partial P_i} {\bf x} \times [\boldsymbol{\sigma}_{ij} + \boldsymbol{\sigma}_{ji}]\boldsymbol{\nu}\,\dif S, \quad i = 1\dots, N.
\end{equation}
where $\boldsymbol{\sigma}_{ij} = \rho^{-1} u_iu_j {\bf I} + \rho(\nabla u_i \cdot \nabla u_j {\bf I} - 2 \nabla u_i \otimes \nabla u_j)$ and
$[\cdot]$ denotes the jump in stress across $\partial \Omega.$ 
\end{proposition}
As described in Specific Aim 3, we solve \eqref{SL1}--\eqref{SL3}  using a double layer potential representation.
While accurate in the far-field, this representation involves a strongly singular integral which
leads to large numerical errors when evaluating the the hydrophobic stress \eqref{stress} along the particle boundaries.
The identity in Proposition \ref{prop:recip} gets around this critical problem by utilizing the fact that $u_j$, $j\neq i$ is smooth and the normal derivative
of  $u_i$ continuous across $\partial P_i$.
